Open archaeology, open source?

### 2024-03-14 analysis/_01-introduction.qmd
Started implementing changes posed by the peer-reviewer.

In their seminal vision for open archaeology, Beck and Neylon [-@beck2012 480-481] identified the movement as comprising a series of principles and practices "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity". Open archaeology therefore promotes more thoughtful scholarly communication practices, most notably in operations relating to publishing, data-sharing, education, and review processes. Archaeologists are also actively engaged in open source software development as means of sharing their research processes and creating tools and resources for general use.

The relationship between open science and open source is complicated by rhetorical claims that have a questionable connection to how academics actually _do_ open source. 
Does academic open source actually make research processes more transparent and improve research outcomes? 
Is it actually boosting efficiency by establishing a common store of knowledge and productive code? 
Is it actually helping to foster new globe-spanning connections and lead to novel research trajectories that would not otherwise come to pass?
Basically, is there more to it than just uploading code and data files to the internet?

Despite hopeful aspirations espoused by open archaeology advocates [see @kansa2014; @kintigh2015], these outcomes are not a given. As per @nguyen2022, @pownall2023 and @leonelli2023 we believe that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results. However, practical circumstances and systemic value regimes that frame what it means to work as an archaeologist presently inhibit the potential for radical transformation, even among open science's most ardent supporters.

@@

This article presents a survey of archaeological software development with two goals in mind:

>>>>

In a 2012 special issue of World Archaeology, Beck and Neylon [-@beck2012 480-481] idenfified open archaeology as comprising a series of principles and practices "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity".
They situate open archaeology as a disciplinary branch of the open science movement, which promotes more thoughtful scholarly communication practices, most notably in operations relating to publishing, data-sharing, education, and review processes.
@ducke2012 similarly argued that archaeologists should more actively engage in open source software development, which involves making source code available for anyone to freely access, modify or reuse, and which complements a common vision of open science as a communal, community-driven, self-correcting and wholly-transparent mode of production.

However, the relationship between open science and open source is complicated by rhetorical claims that have a questionable connection to how academics actually _do_ open source. 
Does academic open source actually make research processes more transparent and improve research outcomes? 
Is it actually boosting efficiency by establishing a common store of knowledge and productive code? 
Is it actually helping to foster new globe-spanning connections and lead to novel research trajectories that would not otherwise come to pass?
Basically, is there more to it than just uploading code and data files to the internet?

Despite these hopeful aspirations espoused by open archaeology advocates [see @kansa2014; @kintigh2015], these beneficial outcomes are not a given.
As per @nguyen2022, @pownall2023 and @leonelli2023 we believe that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
However, practical circumstances and systemic value regimes that frame what it means to work as an archaeologist presently inhibit the potential for radical transformation, even among open science's most ardent supporters.

@@

This article presents a survey of open source archaeological software development with two goals in mind:

### 2024-03-14 analysis/_02-background.qmd
Started implementing changes posed by the peer-reviewer.

Open source is a software development model that prioritizes transparent work processes. Initially driven by the idea that computer users should be free to understand and manipulate the software that they install on their computers (e.g., "free software", as initially conceived by the Free Software Foundation), open source has become a means of collaborative software development [@kelty2008: chapter 3, especially starting at page 99]. By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base. Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision [see @kelty2008 on the emacs saga for a quintessential example]. An open code base may also be used to support alternative projects whose missions diverge from the original plan, and an entire project may be "forked", or taken in a new direction if contributors are dissatisfied with how core developers run things.

@@

However, this is not the same as saying that open source is completely anarchic or based on the "wisdom of crowds". In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants [@oneil2009]. They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term. Open source is more than just putting your code online; to be successful, it requires participation in a social experience [@ratto2003; @kelty2008].

@@

The idea that scientists should generally contribute to a public domain of knowledge without profit motive has led to open science being heralded as revolutionary, community-oriented, and anti-capitalist means of production. However, while open science does have the _potential_ to effect radical change, this is not a given. The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals [@mirowski2018]. Moreover, the open science movement, which is dominated by STEM disciplines, prioritizes a grossly simplified and asocial notion of what science is and entails. Namely, it considers science as the accumulation and assembly of a species-level understanding of the world, which is not held by any one individual but is stored in seemingly value-neutral and disembodied media, facts and observations. This is manifested by digital telecommunications systems that host files, document processes, facilitate co-working opportunities, and perform automated processes. However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science.

Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain [@tennant2020]. But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce. Whereas open source emerged from concern of consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible. Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions. 
Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs [@kling2003]. Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of resume-building, tenure and promotion [@mirowski2018; @dorta-gonzalez2021]. Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out [@carver2022; @adema2021]. Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours [@kelty2008: 271-275; @howison2013]. Open source projects, on the other hand, are motivated by a more practical need for the software to function properly in perpetuity, and contributors may remain actively or sporadically involved to satisfy users' needs, or to direct users to derivative and functional forks of abandoned software [@kelty2008: 278-281; @coleman2012: 116-122; @hippel2003].

@@

Software forges—collaborative web platforms like GitHub, GitLab and Codeberg—are designed to facilitate open source software development by hosting public git repositories.
However, they also support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
These platforms also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, "star" certain repositories as a combine bookmarking and 'like' feature, and maintain a public-facing profile that includes personally-identifying information (e.g. profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.
Code-sharing platforms thus serve as comprehensive developer portfolios and community networking resources.
While these additional features are meant to complement and enhance the experience of contributing to open source projects, they are not actually part of the git protocol.

>>>>

Open source is a software development model that prioritizes transparent work processes. Initially driven by the idea that computer users should be free to understand and manipulate the software that they install on their computers (e.g., "free software", as initially conceived by the Free Software Foundation), open source has become a means of collaborative software development [@kelty2008: chapter 3, especially starting at page 99]. By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base. Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision. An open code base may also be used to support alternative projects whose missions diverge from the original plan, and an entire project may be "forked", or taken in a new direction if contributors are dissatisfied with how core developers run things.

@@

However, this is not the same as saying that open source is completely anarchic or based on the "wisdom of crowds". In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms of social mediation to help delegate and vet contributions made by distributed participants [@oneil2009]. They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term. Open source is more than just putting your code online; to be successful, it requires participation in a social experience [@ratto2003; @kelty2008].

@@

The idea that scientists should generally contribute to a public domain of knowledge without profit motive has led to open science being heralded as revolutionary, community-oriented, and anti-capitalist means of production. However, while open science does have the _potential_ to effect radical change, this is not a given. The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals [@mirowski2018]. Moreover, the open science movement, which is dominated by STEM disciplines, prioritizes a grossly simplified and asocial notion of what science is and entails. Namely, it considers science as the accumulation and assembly of a species-level understanding of the world, which is not held by any one individual but is stored in seemingly value-neutral and disembodied media, facts and observations. This is manifested by information and communication technologies that host files, document processes, facilitate co-working opportunities, and perform automated processes (e.g. digital repositories, reproducible notebooks, automated evaluation of research findings, systems for ensuring stable references to consistent identifiers, etc). However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science.

Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain [@tennant2020]. But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce. Whereas open source emerged from concerns over consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible. Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions. 
Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs [@kling2003]. Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of resume-building, tenure and promotion [@mirowski2018; @dorta-gonzalez2021]. Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out [@carver2022; @adema2021]. Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours [@kelty2008: 271-275; @howison2013]. Scientists may archive their work in an institutional repository, at which point the work enters a stasis state that invites reference to the completed work, but which precludes any potential for the work to be updated, modified, or directly built upon. Open source projects, on the other hand, are motivated by a more practical need for the software to function properly in perpetuity, and contributors may remain actively or sporadically involved to satisfy users' needs, or to direct users to derivative and functional forks of abandoned software [@kelty2008: 278-281; @coleman2012: 116-122; @hippel2003].

@@

Software forges—collaborative web platforms like GitHub, GitLab and Codeberg—are designed to facilitate open source software development by hosting public git repositories.
However, they also support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.^[Much of this relies on proprietary software owned and managed by commercial organisations, and there has been some controversy about the take-over of open source platforms by for-profit entities (see XXXX)].
These platforms also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, "star" certain repositories as a combine bookmarking and 'like' feature, and maintain a public-facing profile that includes personally-identifying information (e.g. profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.
Code-sharing platforms thus serve as comprehensive developer portfolios and community networking resources.
While these additional features are meant to complement and enhance the experience of contributing to open source projects, they are not actually part of the git protocol.

### 2024-03-14 analysis/_03-data_methodology.qmd
Updating reference to the compendium
Started implementing changes posed by the peer-reviewer.

We present an exploratory quantitative analysis of open-archaeo [[open-archaeo.info](open-archaeo.info), @batist2023], a directory of `r nrow(oarch)` pieces of open source archaeological software and other digital resources maintained primarily by one of us (ZB) since 2018.

@@

Open-archaeo is a relatively comprehensive list.
While our initial intention was to only list open source software, its scope has expanded to include all software created by and for archaeologists.
Apart from regular updates by its primary maintainer (ZB), it has been expanded by a wider network of contributors and has benefited from the wider range of domain specialisms this has brought.
However, open-archaeo generally lacks software written before archaeologists started using collaborative software development platforms such as GitHub, and software that is not shared on the web at all.
The dataset is also limited by the experiences of its primary maintainers.^[We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps. Instructions for contributing to open-archaeo can be found at <https://github.com/zackbatist/open-archaeo>.]

@@

Where applicable, we obtained more detailed information about each repository's contents and contribution histories from the GitHub API (application programming interface).
Our analysis incorporates data on `r n_repos` repositories, comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.

We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).
This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or host it elsewhere (`r percent(p_no_github)` of the total), are excluded from these parts of the analysis, though we were still able to perform an analysis of their contents and authorship from the data compiled in open-archaeo itself.
We also cannot include collaboration through offline or private channels, or forms of collaboration we do not know about. 
We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves. 
Our earliest data is from 2005 and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it [@ducke2013; @whallon1972].

@@

The quantitative analyses and figures presented here were generated with `r version$version.string` [@rcoreteam2023].
The full data and code is available in the compendium that accompanies this paper [@roe2023] <!-- TODO: reference should be updated to reflect the actual version number and date -->.

>>>>

We present an exploratory quantitative analysis of open-archaeo [open-archaeo.info, @batist2023], a directory of `r nrow(oarch)` pieces of open source archaeological software and other digital resources maintained primarily by one of us (ZB) since 2018.

@@

Open-archaeo is a relatively comprehensive list.
While our initial intention was to only list open source software, its scope has expanded to include all software created by and for archaeologists.
Apart from regular updates by its primary maintainer (ZB), it has been expanded by a wider network of contributors and has benefited from the wider range of domain specialisms this has brought.
However, open-archaeo generally lacks software written before archaeologists started using collaborative software development platforms such as GitHub, and software that is not shared on the web at all.
It also includes numerous non-software resources, as well as software developed and distributed without the use of software forges.
The dataset is also limited by the experiences of its primary maintainers.^[We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps. Instructions for contributing to open-archaeo can be found at <https://github.com/zackbatist/open-archaeo>.]

@@

Where applicable, we obtained more detailed information about each repository's contents and contribution histories from the GitHub API (application programming interface).
Our analysis incorporates data on `r n_repos` repositories^[We excluded three GitHub repositories from the analysis for technical reasons. For example, https://github.com/carpentries-incubator/R-archaeology-lesson is a repository within the scope of open-archaeo, but it was forked from a pre-existing repository that is not (https://github.com/datacarpentry/R-ecology-lesson) and thus includes in its commit history irrelevant data from the parent repository.], comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.

We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).
This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or which develop and host code on platforms other than GitHub (`r percent(p_no_github)` of the total), were excluded from parts of the analysis that examine *how* archaelogists develop software.
However, we were still able to draw from all records to ascertain the general composition of open-archaeo, and by extension, to address *what* kinds of software and resources archaeologists make.

That being said, we cannot account for practices that occur through offline or private channels, or forms of collaboration we do not know about. 
We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves. 
Our earliest data is from 2005 and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it [@ducke2013; @whallon1972].

@@

The quantitative analyses and figures presented here were generated with `r version$version.string` [@rcoreteam2023].
The full data and code is available in the compendium that accompanies this paper [@roe2024].

### 2024-03-14 analysis/_04-open_archaeology.qmd
Started implementing changes posed by the peer-reviewer.

`r percent(p_platform_r)` of projects are extensions to the statistical programming language R, making it the most widely-used platform by a large margin (@tbl-platforms).
Python, another programming language, is also relatively popular (`r percent(p_platform_python)`), as are plugins for the open source geographic information system QGIS (`r percent(p_platform_qgis)`).
Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g. AutoCAD, ArcGIS), a number of lesser used programming languages, and a genre consisting of custom forms and spreadsheet templates.
Many of these are targeted by only one or two developers; the larger platforms tend to be more diverse.

@@

When we compare categories with thematic tags, we see the general domains that each kind of resource is designed to serve.
We see that packages are fairly common across the board.
Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.
These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.

```{r tbl-tags}
#| tbl-cap: Themes of open archaeology projects
oarch |>
  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
               values_drop_na = TRUE) |>
  drop_na(tag) |>
  group_by(tag, category) |>
  summarise(n = n(), .groups = "drop_last") |>
  mutate(total = sum(n)) |>
  pivot_wider(names_from = category, values_from = n) |>
  mutate(
    tag = glue("[{tag}](https://open-archaeo.info/tags/{slugify(tag)})"),
    across(everything(), \(x) replace_na(x, 0))
  ) |>
  arrange(-total) |>
  rename(theme = tag) |>
  rename_with(tools::toTitleCase) |>
  tt() |>
  format_tt(1, markdown = TRUE)
```

@@

However, since around 2022, the number of new GitHub repositories has significantly plateaud, while the cumulative number of commits has continued to rise.
This may relate to a general emphasis on maintaining existing code and working on established projects, rather than spinning up new ones.

>>>>

`r percent(p_platform_r)` of projects are extensions to the statistical programming language R, making it the most widely-used platform by a large margin (@tbl-platforms).
Python, another programming language, is also relatively popular (`r percent(p_platform_python)`), as are plugins for the open source geographic information system QGIS (`r percent(p_platform_qgis)`).^[While Blender and QGIS plugins are written using the Python language, our intent while categorizing platforms was to get a sense of the developer ecosystems in which archaeological software engineers participate, rather than to simply gauge the popularity of different languages [@batist2023 2].]
Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g. AutoCAD, ArcGIS), a number of lesser used programming languages, and a genre consisting of custom forms and spreadsheet templates.
Many of these are targeted by only one or two developers; the larger platforms tend to be more diverse.

@@

When we compare categories with thematic tags, we see the general domains that each kind of resource is designed to serve.
We see that packages are fairly common across the board.
Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.
These tools may require greater access to system resources, or may require more complex user interfaces than what R or Python IDEs (integrated development environments) tend to provide.

```{r tbl-tags}
#| tbl-cap: Themes of open archaeology projects. See https://github.com/zackbatist/open-archaeo/blob/master/tags.md for a more comprehensive description of each tag's scope.
oarch |>
  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
               values_drop_na = TRUE) |>
  drop_na(tag) |>
  group_by(tag, category) |>
  summarise(n = n(), .groups = "drop_last") |>
  mutate(total = sum(n)) |>
  pivot_wider(names_from = category, values_from = n) |>
  mutate(
    tag = glue("[{tag}](https://open-archaeo.info/tags/{slugify(tag)})"),
    across(everything(), \(x) replace_na(x, 0))
  ) |>
  arrange(-total) |>
  rename(theme = tag) |>
  rename_with(tools::toTitleCase) |>
  tt() |>
  format_tt(1, markdown = TRUE)
```

@@

However, since around 2022, the number of new GitHub repositories has significantly plateaued, while the cumulative number of commits has continued to rise.
This may relate to a general emphasis on maintaining existing code and working on established projects, rather than spinning up new ones.

### 2024-03-14 analysis/_05-collaboration.qmd
Started implementing changes posed by the peer-reviewer.

Perhaps unsurprisingly, given the low uptake of GitHub's collaborative features, `r percent(p_one_contrib)` of open-archaeo repositories only contain commits from a single user.
Even in the minority of projects that have more than one contributor, work (as measured by number of recorded commits) is distributed highly unevenly (@fig-contributions).
The lead maintainer almost always does the lion share of the work: they are responsible for more than half of commits in `r percent(p_big_boy_50)` of projects and more than three-quarters in `r percent(p_big_boy_75)`.
This may be attributed to the steep learning curve commonly attributed to working with git.
While git can be a great way to track changes and manage distributed contributions to a common code base, it can also be unwieldy in situations when multiple users (especially those with less experience using git for collaborative purposes) are expected to contribute within short spans of time.
This meshes with our prior observations that projects which tend to exhibiting higher commit rates have less contributors.
Additionally, our analyses neglects to account for contributions that are not tracked via git or GitHub.
Those who do not code may provide creative guidance or feedback during in-person meetings, via email, or using alternative online messaging or social media platforms.
A more focused qualitative assessment of these non-coding and supportive work practices would shed more light on the totality of effort that goes into producing and maintaining open source projects.

>>>>

Perhaps unsurprisingly, given the low uptake of GitHub's collaborative features, `r percent(p_one_contrib)` of open-archaeo repositories only contain commits from a single user.
Even in the minority of projects that have more than one contributor, work (as measured by number of recorded commits) is distributed highly unevenly (@fig-contributions).
The lead maintainer almost always does the lion share of the work: they are responsible for more than half of commits in `r percent(p_big_boy_50)` of projects and more than three-quarters in `r percent(p_big_boy_75)`.
This may be attributed to the steep learning curve commonly attributed to working with git.
While git can be a great way to track changes and manage distributed contributions to a common code base, it can also be unwieldy in situations when multiple users (especially those with less experience using git for collaborative purposes) are expected to contribute within short spans of time.
This meshes with our prior observations that projects exhibiting higher commit rates have less contributors.
Additionally, our analyses neglect to account for contributions that are not tracked via git or GitHub.
Those who do not code may provide creative guidance or feedback during in-person meetings, via email, or using alternative online messaging or social media platforms.
A more focused qualitative assessment of these non-coding and supportive work practices would shed more light on the totality of effort that goes into producing and maintaining open source projects.

### 2024-03-14 analysis/_06-network.qmd
Started implementing changes posed by the peer-reviewer.

Despite being a minority language, Blender packages are more central than all other package platforms on average, but this is a statistical anomaly caused by uneven sampling.
R is naturally the platform with the next highest average centrality since it serves as a _lingua franca_ that draws developers from across the discipline.
Many of the QGIS plugins add various specialized features to the extensible GIS platform, and are therefore developed by interdisciplinary teams, which explains its high rank.
Python projects, which tend to be infrastructural or are of interest to members of other fields, are also highly ranked in terms of average centrality.
^[While Blender and QGIS plugins are written using the Python language, our intent while categorizing platforms was to get a sense of the developer ecosystems in which archaeological software engineers participate, rather than to simply gauge the popularity of different languages [@batist2023 2].]
Our findings support the notion that there are significant development patterns when working across different languages and platforms, and further analysis to qualify these observation is warranted.

>>>>

Despite being a minority language, Blender packages are more central than all other package platforms on average, but this is a statistical anomaly caused by uneven sampling.
R is naturally the platform with the next highest average centrality since it serves as a _lingua franca_ that draws developers from across the discipline.
Many of the QGIS plugins add various specialized features to the extensible GIS platform, and are therefore developed by interdisciplinary teams, which explains its high rank.
Python projects, which tend to be infrastructural or are of interest to members of other fields, are also highly ranked in terms of average centrality.
Our findings support the notion that there are significant development patterns when working across different languages and platforms, and further analysis to qualify these observation is warranted.
